Simulation of angular resolution of a new electromagnetic sampling calorimeter

### 2022-08-11 paper.tex
u

\setcounter{page}{0}
\title{Simulation of Angular Resolution of a New Electromagnetic Sampling Calorimeter}

@@

This study describes a machine-learning approach based on XGBoost (XGB), which is an optimized distributed gradient-boosting library~\cite{xgboost:2016}. With machine-learning algorithms under the gradient boosting framework, we can deduce the direction of the incident photons entering a sampling calorimeter with alternating layers of Pb plate and plastic scintillator strips in the horizontal and vertical directions.

\section{Electromagnetic Shower Simulation}
\label{sec:ems}

\begin{figure}[!hbt]
\centering
\includegraphics[width=0.6\textwidth]{figures/Fig1_detector_schematic.jpeg}
\caption{ Schematic of the sampling calorimeter model consisting of 105 alternating layers of a Pb and segmented scintillator plate in 35 strips. Each scintillator plate is oriented alternatively in horizontal and vertical directions. See text for details. }
\label{fig:det_conf}
\end{figure}

The sampling calorimeter was designed as a block consisting of alternating layers of an 1-mm-thick Pb absorber and a 5-mm-thick polyvinyltoluene-based plastic scintillator. The plastic scintillator is segmented into 15-mm-wide strips, which are alternatively oriented in the vertical and horizontal directions, as shown in Figure~\ref{fig:det_conf}. It has a cross-section of size 525~$\times$~525~mm$^{2}$ and accommodates 105 alternating layers of 630~mm length, which corresponds to 20 radiation lengths (20$X_{0}$) that are sufficiently long to absorb full photon energy in the range of 0.1 to 2~GeV.

@@

\section{Reconstruction of Incident Angles}
\label{sec:res}

@@

Angular resolutions were deduced by varying the strip width from 5 to 35 mm, as shown in Figure~\ref{fig:angle_reco_width}. The narrower the strip width, the larger is the size of the features. Consequently, the angle-reconstruction performance declines. However, the wider strips hardly provide detailed information on the EM shower. The 15-mm-wide strips yield a good angular resolution of 1.23~$\pm$~0.01$^{\circ}$.

@@

Focusing only on the angle reconstruction, we designed the detector with the smallest number of layers. Figure~\ref{fig:angle_reco_layer} (a) shows the angles reconstructed using the front 32 layers of the detector, which correspond to 6.2$X_{0}$, for 1-GeV photons. A fraction of the photons fail to be reconstructed because of insufficient information gathering caused by deeply penetrating photons without shower generation. The failed events are represented as a delta function near 0, and such events are removed by fixing the total energy deposit at a value higher than 10~MeV, which is 1\% of the incident energy. The angular resolution of the front layer is estimated to be 1.30~$\pm$~0.01$^{\circ}$ when 1.9\% of the photons fail to be reconstructed.

@@

\begin{figure}[!hbt]
\centering
\includegraphics[width=0.44\textwidth]{figures/Fig7_p0.pdf}
\includegraphics[width=0.44\textwidth]{figures/Fig7_p1.pdf}
\caption{ $p_{0}$ and $p_{1}$ as functions of the incident angle }
\label{fig:res_edep}
\end{figure}

Figure~\ref{fig:res_edep} shows the estimated $p_{0}$ and $p_{1}$ as functions of the incident angle. The average values of $p_{0}$ and $p_{1}$ are estimated to be 0.24$^{\circ}$ and 1.24$^{\circ}$, respectively, for $\theta>$~10$^{\circ}$. For a small $\theta$, $p_{1}$ is smaller than that for a larger $\theta$ because the angular resolution is largely dependent on the incident angle for low incident energies, whereas the opposite trend can be observed for $p_{0}$.

@@

The angular resolution depends on the incident energy of the photon, and the dependence is expressed as $p_{0} \oplus p_{1}/\sqrt{E_{\gamma}}$. Except for small incident angles ($\theta<$~15~deg), the two parameters do not significantly change up to 
$40^{\circ}$, and the dependence is estimated to be 0.24 $\oplus$ 1.24$^{\circ}/\sqrt{E_{\gamma}}$.

>>>>

\setcounter{page}{0}
\title{Simulation of angular resolution of a new electromagnetic sampling calorimeter}

@@

This study describes a machine-learning approach based on XGBoost (XGB), which is an optimized distributed gradient-boosting library~\cite{xgboost:2016}. With machine-learning algorithms under the gradient-boosting framework, we can deduce the direction of the incident photons entering a sampling calorimeter with alternating layers of Pb plate and plastic scintillator strips in the horizontal and vertical directions.

\section{Electromagnetic shower simulation}
\label{sec:ems}

\begin{figure}[!hbt]
\centering
\includegraphics[width=0.6\textwidth]{figures/Fig1_detector_schematic.jpeg}
\caption{ Schematic of the sampling calorimeter model consisting of 105 alternating layers of a Pb and segmented scintillator plate in 35 strips. Each scintillator plate is oriented alternately in horizontal and vertical directions. See text for details. }
\label{fig:det_conf}
\end{figure}

The sampling calorimeter was designed as a block consisting of alternating layers of an 1-mm-thick Pb absorber and a 5-mm-thick polyvinyltoluene-based plastic scintillator. The plastic scintillator is segmented into 15-mm-wide strips, which are alternately oriented in the vertical and horizontal directions, as shown in Figure~\ref{fig:det_conf}. It has a cross-section of size 525~$\times$~525~mm$^{2}$ and accommodates 105 alternating layers of 630~mm length, which corresponds to 20 radiation lengths (20$X_{0}$) that are sufficiently long to absorb full photon energy in the range of 0.1 to 2~GeV.

@@

\section{Reconstruction of incident angles}
\label{sec:res}

@@

Angular resolutions were deduced by varying the strip width from 5 to 35 mm, as shown in Figure~\ref{fig:angle_reco_width}. The narrower the strip width, the larger is the size of the features. Consequently, the angle-reconstruction performance is better with 10-mm-wide strips than 5-mm-wide strips. However, the wider strips hardly provide detailed information on the EM shower. The 15-mm-wide strips yield a good angular resolution of 1.23~$\pm$~0.01$^{\circ}$.

@@

Focusing only on the angle reconstruction, we designed the detector with the smallest number of layers. Figure~\ref{fig:angle_reco_layer} (a) shows the angles reconstructed using the front 32 layers of the detector, which correspond to 6.2$X_{0}$, for 1-GeV photons. A fraction of the photons fail to be reconstructed because of insufficient information gathering caused by deeply penetrating photons without shower generation. The failed events are represented as a delta function near 0, and such events are removed by requiring the total energy deposit to be higher than 10~MeV, which is 1\% of the incident energy. The angular resolution of the front layer is estimated to be 1.30~$\pm$~0.01$^{\circ}$ when 1.9\% of the photons fail to be reconstructed.

@@

\begin{figure}[!hbt]
\centering
\includegraphics[width=0.44\textwidth]{figures/Fig7_p0.pdf}
\includegraphics[width=0.44\textwidth]{figures/Fig7_p1.pdf}
\caption{ $p_{0}$ and $p_{1}$ as a function of the incident angle }
\label{fig:res_edep}
\end{figure}

Figure~\ref{fig:res_edep} shows the estimated $p_{0}$ and $p_{1}$ as a function of the incident angle. The average values of $p_{0}$ and $p_{1}$ are estimated to be 0.24$^{\circ}$ and 1.24$^{\circ}$, respectively, for $\theta>$~10$^{\circ}$. For a small $\theta$, $p_{1}$ is smaller than that for a larger $\theta$ because the angular resolution is largely dependent on the incident angle for low incident energies, whereas the opposite trend can be observed for $p_{0}$.

@@

The angular resolution depends on the incident energy of the photon, and the dependence is expressed as $p_{0} \oplus p_{1}/\sqrt{E_{\gamma}}$. Except for small incident angles ($\theta<$~$^{\circ}$~deg), the two parameters do not significantly change up to 
$40^{\circ}$, and the dependence is estimated to be 0.24 $\oplus$ 1.24$^{\circ}/\sqrt{E_{\gamma}}$.
